### 2015-06-09 text/ms.Rmd
Major rewrite of introduction and methods

When models give conflicting estimates of a population's status, there are a
number of solutions. (1) We can pick a single model. However, it isn't always
obvious which model is best, and even objectively second-best models may contain
useful additional information. (2) We can maintain multiple models throughout
the decision-making process. However, this can be complicated (REFs)... (3) We
can combine the model outputs. This could be as simple as taking the average,
potentially weighting the models by some performance metric. This could be as
complicated as forming an additional 'hyper' model that draws on interactions
between model outputs and potentially incorporates additional information to
derive a best estimate.

Ensemble models are widely used in the climate sciences, where thousands of
models with varying structural combined across different structural ... The
idea behind ensemble models forms the backbone of many machine learning
methods. For example...

In fisheries science, a common problem is estimating the status of an exploited
fish population. For the majority of fish stocks, we have limited information to
go on and stock status isn't known. In recent years, a number of methods have
been proposed to derive population status based on limited information and a set
of assumptions. However, we know that these models can give conflicting output
and no one model is that accurate.

Motivation: (Fig. \ref{fig:motivate}).

Ensemble models combine estimates of individual models, potentially along with
additional covariates, in an additional 'ensemble' model
(Fig. \ref{fig:didactic}).

Exploit interactions between individual models and relationships between other
conditions of the data and the performance of the individual models.

Fitted to populations of known or assumed status and this ensemble model is
used to extrapolate to populations of interest.

Here, we develop ensemble models for data-limited exploited fish populations. We
explore a variety of ensemble approaches applied to both simulated and
real-world fish stocks and compare performance...

@@

We developed and tested ensemble models with both a fully factorial simulated
dataset [@rosenberg2014] and the RAM Legacy Stock Assessment Database
[@ricard2012]. @rosenberg2014 provide a full description of the simulation
model and the code to generate the simulations is available at
<https://github.com/flr/StockSims> (TODO: not there). To summarize, the models were built in the
package FLR [@kell2007] for the statistical software \textsf{R} [@r2015]...
included all combinations of the following factors:

- Three life histories: small pelagic, demersal, large pelagic.
- Two time series lengths: 20 and 60 years.
- Three levels of initial depletion: 1, 0.7, and 0.4.
- Four harvest dynamics: (1) a constant harvest rate, (2) a harvest rate that
  is coupled with biomass to mimic an open-access single-species fishery, (3) a
  one-way trip where harvest rate increases five percent per year to 80% of the
  level at which the stock would crash, and (4) roller-coaster shaped where the
  harvest rate increases, maintains a steady level at 80% of the level at which
  the stock would crash, and then decreases to FMSY. 
- Two levels of multiplicative standard deviation around catch (0 and 0.2)
- Two levels of standard deviation of recruitment variability (0.2 and 0.6)
- Two levels of autoregressive correlation on recruitment variability (0 and 0.6)

\noindent @rosenberg2014 ran ten iterations for each combination of factors
adding stochastic draws of recruitment and catch-recording variability each
time to generate a total of 5760 stocks.

Our analysis of the RAM Legacy Stock Assessment Database was based on version
XX downloaded on XX. This version includes XX stocks from XX countries across
XX taxonomic orders.

@@

We fit four individual data-limited models to estimate \bbmsy\\. Three of the
models are mechanistic and based generally on Shaefer biomass dynamics
[@schaefer1954] of the form:

@@

\noindent where $B_t$ represents biomass at time $t$, $r$ represents the
intrinsic population growth rate, $B_0$ represents unfished biomass, and $C$
represents catch. One of the models is empirical instead of mechanistic.
@rosenberg2014 provide a full summary of these four methods and we provide some
further details in the Supporting Materials. Code to carry out all methods is
available in an accompanying package **datalimited** for the statistical
software \textsf{R}. In summary:

@@

- *mPRM* (modified panel regression model) is a modified version of the
panel-regression model used in @costello2012. Unlike the other models
considered, mPRM is empirical and not mechanistic --- it fits a linear model to
a series of characteristics of the catch time series and stock with \bbmsy\\ as
the response. The model used in this paper and @rosenberg2014 is modified from
the original in that it simplifies the life-history categories to be applicable
to simulated data.

@@

Ensemble methods allow us to incorporate additional covariates into our models
and potentially leverage interactions of how these covariates combine with the
main individual model estimates to estimate population status. Examples might
be life-history characteristics, additional information on exploitation
patterns, or additional statistical properties of the data going into the
individual models. For simplicity, and to allow us to apply models developed
with the simulated dataset to the real-world dataset, we added only one set of
additional covariates: spectral properties of the catch time series itself. We
fit spectral models to the catch time series and recorded representative
short-term and long-term spectral densities of 0.05 (corresponding to a 20-year
cycle) and 0.2 (corresponding to a 5-year cycle).

@@

The individual models we seek to combine with ensemble models provide time
series of \bbmsy\\. Therefore, we can use ensemble models to estimate a number of
different attributes of the time series. Here, we focus on two attributes the
mean and slope of \bbmsy\\ in the last 3 years. We refer to these quantities
collectively as $\theta$. Together, these quantities address both the current
state and trajectory of status, both of which may be of management and
conservation interest. To avoid undue influence of the end points of the time
series on the calculated slope, we instead measured the median slope as the
Theil-Sen estimator [@theil1950].

An ensemble model can combine inference through anything from a simple average,
to a weighted average, to a complex non-linear model. Here we choose four
ensemble models to compare: an average, a linear model with two-way
interactions, a random forest, and boosted regression tree.

General form:

\begin{equation}
\hat{\theta} = \mathrm{argmin}_\theta(L(b \, | \, \hat{b}_{\mathrm{ensemble}}))
\end{equation}

\begin{equation}
\hat{b}_\mathrm{ensemble} = f(\hat{b}_\mathrm{CMSY}, \hat{b}_\mathrm{COM-SIR}, \hat{b}_\mathrm{SSCOM}, \hat{b}_\mathrm{mPRM}, \mathrm{other~covariates} \, | \, \theta)
\end{equation}

\noindent where $L$ is a generalized loss function, $f$ is some generalized regression function, $\hat{b}_\mathrm{COM-SIR}$ is the prediction from COM-SIR, etc., and $\theta$ is the set of parameters for the regression function.

*Linear model*: For populations $i$ $1$ through $n$, individual models and
covariates $j$ $1$ through $m$, and allowing $\theta$ to refer to both
quantities estimated from individual models and additional covariates, we fit
the linear model with two-way interactions as:

@@

\begin{equation}
\theta_{\mathrm{ensemble}(i)} = 
  \beta_0 + 
  \sum_{j=1}^m \beta_j \theta_{j(i)} + 
  \sum_{j=1}^m \sum_{\substack{k =1 \\ j \neq k}}^m 
            \beta_{kj} \theta_{k(i)} \theta_{j(i)} +
  \epsilon_i, \quad
  \epsilon_i \sim \mathrm{Normal}(0, \sigma),
\end{equation}

\noindent
where the $\beta$ represent coefficients and $\epsilon$ represents normally
distributed residual variation. We also tried fitting linear models without
interactions but found better cross-validation predictive performance from all
performance measures evaluated using the procedure outlined below.

*Random forests* are a machine learning method that combines a series of
regression trees. Regression trees sequentially determine what value of
a predictor best splits the response data into two 'branches' based on
a specified loss function. In random forests, a series of regression trees are
built on a random subset of the data and random subset of the covariates of the
model. By combining many of these stochastically generated models, random
forests can provide good predictive performance and fit highly non-linear
relationships TODO. We fit random forest models with the **randomForest**
package [@liaw2002] for \textsf{R} with the default argument values.

*Boosted regression models* are a machine learning method that, like random
forests, combines a series of regression trees built on random subsets of the
data, but also add a boosting procedure. Boosting refers to fitting a series
of models with each subsequent model fit to the residuals from the previous
model; data points that are fit poorly in a given model are given more weight
in the next model [@elith2008]. Boosting can result in TODO and TODO and is
a common and powerful machine learning technique [@elith2008]. We fit boosted
regression tree models with the **GBM** package [@ridgeway2015] for \textsf{R}
with 10000 trees, an interaction depth of $2$, a learning rate (shrinkage
parameter) of $0.01$, and all other arguments at their default values.

For all ensemble models of $\overline{B/B_\mathrm{MSY}}$, which is bounded at
zero, we log transformed all $\overline{B/B_\mathrm{MSY}}_{(ji)}$ and
exponentiated the estimated $\log
\widehat{\overline{B/B_\mathrm{MSY}}}_\mathrm{ensemble}$ to derive a median
estimate of $\widehat{\overline{B/B_\mathrm{MSY}}}_\mathrm{ensemble}$. For the
estimates of \bbmsy\\ slope, which are not bounded at zero, we fit ensemble
models on the natural untransformed scale.

@@

A critical component to any predictive modelling exercise is to evaluate the
predictive performance of a model on new data [@hastie2009]. Typically, data
are limited in availability, and so a common and effective tool is
cross-validation [@hastie2009]. We used repeated three-fold cross validation to
test predictive performance: we randomly divided the dataset into three sets,
built ensemble models on two-thirds of the data, and evaluated predictive
performance on the remaining third of the data. We repeated this across each of
the three splits and then repeated the whole procedure XX times to account for
bias that may result from any one set of validation splits.

Predictive performance can be evaluated with metrics that represent a variety
of modelling goals. For instance, a model may provide an unbiased but highly
imprecise prediction. For continuous response variables such as those
considered here performance metrics often measure some form of bias, precision,
or accuracy (a combination of bias and precision), or the ability to correctly
rank or correlate across populations. Here, we measure proportional error (PE;
also called 'relative error') defined as $(\hat{\theta} - \theta)/\theta$,
where $\theta$ represents the true parameter value and $\hat{\theta}$ the
estimate of that parameter ($\overline{B/B_\mathrm{BMSY}}$ or slope of
\bbmsy\\). We summarize median proportional error  to measure bias, median
absolute proportional error to measure accuracy, and Spearman's rank-order
correlation to measure the ability to correctly rank population status.

>>>>

Ensemble models, which combine the output from multiple models, are one
potential solution to this problem. A simple ensemble model might take the
average of multiple model outputs, perhaps weighted by some goodness of fit
(REFs). A complex ensemble model might also take into account the interactions
between model outputs and additional covariates. Ensemble models are popular in
many fields including public health [e.g. @thomson2006], agriculture [e.g.
@cantelaube2005], climate science (e.g. REFs), and machine learning (e.g.
REFs). We think they have tremendous opportunity in ecological resource
management.

In fisheries science, a common problem is estimating the status of an exploited
fish population. For the majority of fish stocks, we have limited biological
information and stock status isn't known (e.g. REFs).  In recent years, a
number of methods have been proposed to derive population status based on the
limited information that is frequently available: fisheries catch time series
and basic knowledge about the productivity of the species. Recent work has
investigated the ability for these models to estimate population status through
a large-scale simulation model [@rosenberg2014]. A key finding was that these
models frequently disagree on population status and any one model has
relatively poor predictive ability (e.g. Fig. \ref{fig:motivate}).

<!--
Exploit interactions between individual models and relationships between other
conditions of the data and the performance of the individual models.
Fitted to populations of known or assumed status and this ensemble model is
used to extrapolate to populations of interest.
-->

Here, we develop ensemble models to estimate the population status of exploited
fish populations. We explore a variety of ensemble approaches applied to both
simulated and real-world fish stocks and compare predictive performance. In
developing these models we highlight the types of problems one encounters when
forming and evaluating ensemble models. We find that ensembles generally
provide improved predictive performance than any single model---especially when
considered across multiple dimensions of predictive ability. Furthermore, the
output from ensemble models can provide better mechanistic understanding of how
invidual models perform.

@@

We first developed and tested ensemble models with a fully factorial simulated
dataset [@rosenberg2014]. The simulation models included three life histories (small
pelagic, demersal, large pelagic), three levels of how depleted biomass was at
the start of the dataset (100%, 70%, and 40% of carrying capacity), four
harvest dynamics (a constant harvest rate, a harvest rate that is coupled with
biomass to mimic an open-access single-species fishery, a scenario where
harvest rate increases to a fixed level, a 'roller-coaster' scenario where the
harvest rate increases, maintains a steady level and decreases to a lower
steady level). Process noise was introduced to the models through two levels of
multiplicative recruitment variability ($N(0, 0.2)$ and $N(0, 0.6)$, two levels
of first-order autoregressive correlation on recruitment variablity (0 and
0.6). The simulation included a scenario without observation error and with
multiplicative observation error around catch at $N(0, 0.2)$.

@rosenberg2014 ran ten iterations for each combination of factors adding
stochastic draws of recruitment and catch-recording variability each time to
generate a total of 5760 stocks. The simulation models were built in the
package FLR [@kell2007] for the statistical software \textsf{R} [@r2015].
@rosenberg2014 provide a full description of the simulation model and the code
to generate the simulations is available at <https://github.com/flr/StockSims>
(TODO: not there).

We also tested our ensemble models on the RAM Legacy Stock Assessment Database
[@ricard2012]. Our analysis of the stock-assessment database was based on
version XX downloaded on XX. After removing stocks for which the models
described below did not converge, this database included `r ram_stocks_n` from
XX countries across XX taxonomic orders.

@@

We fit four individual data-limited models to estimate \bbmsy\\. Three of the
models are mechanistic and based generally on @schaefer1954 biomass
dynamics of the form:

@@

\noindent where $\hat{B}_{t+1}$ represents predicted biomass at time $t$ plus
one year, $B_t$ represents biomass at time $t$, $r$ represents the intrinsic
population growth rate, $B_0$ represents unfished biomass or carrying capacity,
and $C$ represents catch. @rosenberg2014 provide a full summary of these four
methods and code to fit all the models is available in an accompanying package
**datalimited** for the statistical software \textsf{R}. In summary:

@@

- *mPRM* (modified panel regression model) is a modified version of the
panel-regression model used in @costello2012. Unlike the other models
considered, mPRM is empirical and not mechanistic --- it fits a linear model to
a series of characteristics of the catch time series and stock with \bbmsy\\ as
the response. The model used in this paper and @rosenberg2014 is modified from
the original in that it simplifies the life-history categories to be applicable
to simulated data.
<!--TODO: what are we doing with the RAM stocks again?-->

@@

Ensemble methods allow us to incorporate additional covariates into our models
and potentially leverage interactions of how these covariates combine with the
main individual model estimates to estimate population status. Examples might
be life-history characteristics, additional information on exploitation
patterns, or additional statistical properties of the data going into the
individual models. For simplicity, and to allow us to apply models developed
with the simulated dataset to the real-world dataset, we added only one set of
additional covariates: spectral properties of the catch time series itself.
Spectral analysis decomposes a time series into the frequency domain and
provides a useful means of describing the statistical properties of the catch
series that is independent of time series length and magnitude of catch. We fit
spectral models to the scaled catch time series (divided by maximum catch) with
the `spec.ar` function in \textsf{R} and recorded representative short- and
long-term spectral densities of 0.05 and 0.2, which correspond to 20- and
5-year cycles.

@@

The individual models we seek to combine with ensemble models provide time
series of \bbmsy\\ ($\hat{b}_\mathrm{CMSY}$, $\hat{b}_\mathrm{COM-SIR}$, etc.).
Therefore, we can use ensemble models to estimate multiple properties of these
status time series. Here, we focus on two values: the mean and slope of
\bbmsy\\ in the last 5 years. Below, we refer to these quantities collectively
as $\hat{b}_\mathrm{ensemble}$. Together, these quantities address the current
state and trajectory of status, both of which may be of management and
conservation interest. To avoid undue influence of the end points of the time
series on the calculated slope, we measured the slope as the Theil-Sen
estimator of median slope [@theil1950].

An ensemble model can combine models through anything from a simple average to
a complex non-linear model. Here we choose four ensemble models to compare: an
average, a linear model with two-way interactions, a random forest, and boosted
regression tree. The general form of our models is as follows:

$$
\hat{\theta} = \mathrm{argmin}_\theta(L(b \, | \, \hat{b}_{\mathrm{ensemble}}))
$$

$$
\hat{b}_\mathrm{ensemble} = 
f(\hat{b}_\mathrm{CMSY}, \hat{b}_\mathrm{COM-SIR}, 
\hat{b}_\mathrm{SSCOM}, \hat{b}_\mathrm{mPRM}, 
S(0.2), S(0.5), | \, \theta)
$$

\noindent where $L$ is a generalized loss function, $f$ is some generalized
regression function, $\hat{b}_\mathrm{CMSY}$ is the prediction from COM-SIR,
$\hat{b}_\mathrm{COM-SIR}$ etc., $b$ is the true value the model is fitted to,
$S(n)$ represents the spectral density at frequency $n$ of the scaled catch
series, and $\theta$ is the set of parameters for the regression function.

<!--
*Linear model*: For populations $i$ $1$ through $n$, individual models and
covariates $j$ $1$ through $m$, and allowing $\theta$ to refer to both
quantities estimated from individual models and additional covariates, we fit
the linear model with two-way interactions as:
-->

@@

<!--
\begin{equation}
\theta_{\mathrm{ensemble}(i)} = 
  \beta_0 + 
  \sum_{j=1}^m \beta_j \theta_{j(i)} + 
  \sum_{j=1}^m \sum_{\substack{k =1 \\ j \neq k}}^m 
            \beta_{kj} \theta_{k(i)} \theta_{j(i)} +
  \epsilon_i, \quad
  \epsilon_i \sim \mathrm{Normal}(0, \sigma),
\end{equation}

\noindent
where the $\beta$ represent coefficients and $\epsilon$ represents normally
distributed residual variation. We also tried fitting linear models without
interactions but found better cross-validation predictive performance from all
performance measures evaluated using the procedure outlined below.
-->

*Random forests* are a machine learning method that combines a series of
regression trees. Regression trees sequentially determine what value of a
predictor best splits the response data into two 'branches' based on a
specified loss function. In random forests, a series of regression trees are
built on a random subset of the data and random subset of the covariates of the
model. By combining many of these stochastically generated models, random
forests can provide strong predictive performance and fit highly non-linear
relationships (REFs). We fit random forest models with the **randomForest**
package [@liaw2002] for \textsf{R} with the default argument values.

*Boosted regression models* are a machine learning method that, like random
forests, combines a series of regression trees built on random subsets of the
data, but also add a boosting procedure. Boosting refers to fitting a series of
models with each subsequent model fit to the residuals from the previous model;
data points that are fit poorly in a given model are given more weight in the
next model [@elith2008]. Boosting can result in TODO and TODO and is a common
and powerful machine learning technique [@elith2008]. We fit boosted regression
tree models with the **GBM** package [@ridgeway2015] for \textsf{R}. Based on
cross-validation with the **caret** package (Supporting Materials, Fig. SX), we
fit GBMs with 2000 trees, an interaction depth of $6$, a learning rate
(shrinkage parameter) of $0.01$, and all other arguments at their default
values.

For all ensemble models of mean \bbmsy\\, which is bounded at zero, we fit the
models in log space and exponentiated the predictions to derive a median
estimate of \bbmsy\\. For the estimates of \bbmsy\\ slope, which are not
bounded at zero, we fit ensemble models on the natural untransformed scale.

@@

A critical component to any predictive modelling exercise is to evaluate the
performance of a model new data [@hastie2009]. We used repeated three-fold
cross validation to test predictive performance: we randomly divided the
dataset into three sets, built ensemble models on two-thirds of the data, and
evaluated predictive performance on the remaining third of the data. We
repeated this across each of the three splits and then repeated the whole
procedure XX times to account for bias that may result from any one set of
validation splits. Since the dynamics of simulated populations differing only
in the magnitude of observation error around catch, magnitude of recuitment
variability, or stochastic draw of these values are likely similar, we grouped
these characteristics together in the cross-validation process---we never
trained and tested on models differing only in these characteristics.

Predictive performance can be evaluated with metrics that represent a variety
of modelling goals. For continuous response variables such as the mean and
slope of population status, peformance metrics often measure some form of bias,
precision, accuracy (a combination of bias and precision), or the ability to
correctly rank or correlate across populations. Here, we measure proportional
error (PE; also called 'relative error') defined as $(\hat{b} - b)/b$. We
summarize median proportional error to measure bias, median absolute
proportional error to measure accuracy, and Spearman's rank-order correlation
to measure the ability to correctly rank populations.
